---
title: "fpcFeatureEngine -- usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fpcFeatureEngine-usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
    )
```
This vignette demonstrates the usage of the R6 class **fpcFeatureEngine**
in the R package **fpcFeatures**.

More specifically, we demonstrate the following practical use case:

*   Compute the functional principal components (FPC)
    based on a training time series data set.

*   Subsequently, compute the FPC **scores**,
    with respect to the aforementioned FPC's
    (i.e. extracted from training data),
    but for a new collection of time series (i.e. different from training data).

Functional principal components are also commonly referred to as **harmonics**.

### **fpcFeatureEngine** assumes that:

*   each time point in each input time series corresponds to a date, and

*   for each given input time series,
    all of its time points come from the same calendar year.

However, for each pair of input time series for **fpcFeatureEngine**,
their respective collections of time points need not coincide.

# 1. Preparing demo data frames: **DF.training**, **DF.new**

### Load the **IW4** remote sensing time series data frame.

We will use the data frame **IW4**, which is bundled
with the R package **fpcFeatures**.

First, load the **fpcFeatures** R package, and the other required packages
for this vignette:
```{r setup}
require(fpcFeatures)
require(ggplot2)
require(knitr)
require(logger)
```

We can now load the embedded data frame **IW4**, and examine its structure.
```{r}
data(IW4);
str(IW4);
```

The **IW4** data frame contains VV time series measurements from approximately
5500 locations (each specified by geocoordinates) for the growing seasons
of the calendar years 2017, 2018 and 2019.
For each location and each year, the corresponding measurement time series
starts around early April and extends to around late October or early November,
at an observation frequency of one measurement every 12 days.

Here are the first few rows of **IW4**:
```{r headIW4}
knitr::kable(head(IW4), align = "c")
```

### Form the **DF.VV** data frame.

In what follows, we will work with the VV variable.
We form a new data frame **DF.VV** by extracting from **IW4**
the VV measurements, the dates and the geocoordinates.
**fpcFeatureEngine** requires its input data frame
to have its location identifier as a single variable.
We therefore create a custom unique location idenfitier
by concatenating the two geocoordinates:
```{r}
DF.VV <- IW4[,c('x','y','date','VV')];
DF.VV[,'x_y'] <- apply(
    X      = DF.VV[,c('x','y')],
    MARGIN = 1,
    FUN    = function(x) { return(paste(x,collapse="_")) }
    );
```

### Split **DF.VV** -- by location -- into **DF.training** and **DF.new**.

```{r}
locations <- unique(DF.VV[,'x_y']);

set.seed(7654321);
is.subset.1 <- sample(x = c(TRUE,FALSE), size = length(locations), replace = TRUE);
locations.1 <- locations[ is.subset.1];
locations.2 <- locations[!is.subset.1];

DF.training <- DF.VV[DF.VV[,'x_y'] %in% locations.1,];
colnames(DF.training) <- paste0("my_",colnames(DF.training));

DF.new <- DF.VV[DF.VV[,'x_y'] %in% locations.2,];
colnames(DF.new) <- paste0("your_",colnames(DF.new));
```

Here are the first few rows of **DF.training**:
```{r}
knitr::kable(head(DF.training), align = "c")
```
We examine the structure of **DF.training**:
```{r}
str(DF.training)
```
We remark here that only the following three columns
are required to invoke **fpcFeatureEngine**:

*  **my_date**: date (of remote sensing observation), must be of type Date
*  **my_x_y**: location identifier, must be of type character
*  **my_VV**: target variable (remote sensing measurement), must be of type numeric

**fpcFeatureEngine** ignores all other columns.

The first few rows of **DF.new**:
```{r}
knitr::kable(head(DF.new), align = "c")
```
Note that **DF.new** has (of course) the same structure as **DF.training**,
but the two data frames have different column names.
**fpcFeatureEngine** does NOT require the training and new data frames
to have the same column names.
We deliberately give different column names to **DF.training** and **DF.new**
to highlight this.

# 2. Training

Recall that "training" here means:
**compute and store functional principal components based on training data**.

First, we instantiate an **fpcFeatureEngine** object,
with the input (training) data and instantiation parameters:
```{r}
my.fpcFeatureEngine <- fpcFeatureEngine$new(
    training.data       = DF.training,
    location            = 'my_x_y',
    date                = 'my_date',
    variable            = 'my_VV',
    min.date            = NULL,
    max.date            = NULL,
    n.partition         = 100,
    n.order             =   3,
    n.basis             =   9,
    smoothing.parameter =   0.1,
    n.harmonics         =   7
    );
```
__Aside: when to supply non-NULL values for the **min.date** and **max.date** input parameters__
<br>
_The **min.date** and **max.date** parameters in the above call to
**fpcFeatureEngine$new( )** could have been omitted since we merely supplied
their default value (i.e. NULL). We explain here when to supply non-NULL
values for either of these parameters.
When computing FPC scores for new data -- see Section 3 below --
**fpcFeatureEngine** requires that the range of observation dates in the
new data encompasses that of the training data.
This is because,
prior to computing the FPC scores,
with respect to FPC extracted from the training data,
for each new data time series,
a B-spline first needs to be fitted through that new data time series,
and each such B-spline must be well-defined
__over the entire observation date range of the training data__.
Conversely, in practice, if one encounters a scenario where the training data
observation date range does not entirely reside within that of the new data,
then one uses the **min.date** and/or **max.date** parameter when instantiating
**fpcFeatureEngine** in order to restrict the subsequent FPC computations
to the observation date range of the new data._
See Section 5, and the plot therein, for an illustration of this.
<br>
<br>

We can now perform the training
(i.e. computation of the functional principal components, a.k.a. the harmonics,
based on the training data),
simply by using the **fpcFeatureEngine$fit( )** function, as follows:
```{r}
logger::log_threshold(level = logger::ERROR);
my.fpcFeatureEngine$fit();
```
The training is complete.
The **log_threshold( )** function call above is just to suppress
the printing of info messages during the fitting process.

If we wish, we could examine the computed harmonics
(functional principal components)
using the **fpcFeatureEngine$plot.harmonics( )** function:
```{r harmonics, fig.height = 28, fig.width = 16}
my.fpcFeatureEngine$plot.harmonics();
```

# 3. Compute FPC scores for new data using the trained **fpcFeatureEngine** object.

We compute the functional principal component scores for the time series
in **DF.new** using the **fpcFeatureEngine$transform( )** function:
```{r transform}
DF.bspline.fpc <- my.fpcFeatureEngine$transform(
    newdata  = DF.new,
    location = 'your_x_y',
    date     = 'your_date',
    variable = 'your_VV'
    );
```

### Examining the output of **fpcFeatureEngine$transform( )**

We examine the structure of the object **DF.bspline.fpc**
returned by **fpcFeatureEngine$transform( )**:
```{r}
str(DF.bspline.fpc)
```

**DF.bspline.fpc** is simply a data frame,
though with a large number of columns.
Here are all of its column names:
```{r}
colnames(DF.bspline.fpc)
```

*   Each row of **DF.bspline.fpc** corresponds to a (location, year).
    In other words, the two columns **your_x_y** and **year** together serve
    as the primary key of the data frame **DF.bspline.fpc**.

*   The integer column names are "date indexes",
    where 1 refers to New Year's Day, 2 = January 2, and so on.

*   On each row -- i.e. for each (location, year) -- the values
    under the date indexes (those with integer column names)
    are the values of the B-spline fitted through the original time series
    of that (location, year).

*   The values under the columns **fpc_1**, **fpc_2**, ... , **fpc_7**
    are the functional principal component scores.

Here are the first few rows of **DF.bspline.fpc**,
with the B-spline columns suppressed:
```{r}
knitr::kable(head(DF.bspline.fpc[,c('your_x_y','year',paste0('fpc_',1:7))]), align = "c", row.names = FALSE)
```

# 4. Visualize FPC scores of new data.

Recall that we discarded the land cover column
when we formed the **DF.training** and **DF.new** data frames,
since land cover is not needed by **fpcFeatureEngine**,
i.e. land cover is not needed for computing functional principal components
or scores.

But now, for the visualization, we wish to indicate the different land covers
with different colours. So, we attach back the land cover information.

The following code segment extracts the land cover of each location
in the original data frame **IW4**, and stores the information
in the data frame **DF.land.cover**:
```{r}
DF.land.cover <- unique(IW4[,c('x','y','land_cover')])
DF.land.cover[,'x_y'] <- apply(
    X      = DF.land.cover[,c('x','y')],
    MARGIN = 1,
    FUN    = function(x) { return(paste(x,collapse="_")) }
    );
DF.land.cover <- DF.land.cover[,c('x_y','land_cover')]
```

The first few rows of **DF.land.cover**:
```{r}
knitr::kable(head(DF.land.cover), align = "c", row.names = FALSE)
```

We attach the land cover back to **DF.bspline.fpc** simply as follows:
```{r}
DF.bspline.fpc <- merge(
    x    = DF.bspline.fpc,
    y    = DF.land.cover,
    by.x = 'your_x_y',
    by.y = 'x_y'
    )
```

We discard the B-spline columns (no longer needed for rest of this vignette)
from **DF.bspline.fpc**, and store the resulting data frame as **DF.fpc**:
```{r}
DF.fpc <- DF.bspline.fpc[,c('your_x_y','year','land_cover',paste0('fpc_',1:7))]
```

The first few columns of **DF.fpc**:
```{r}
knitr::kable(head(DF.fpc), align = "c", row.names = FALSE)
```

The following code segment generates the scatter plot of
**fpc_1** against **fpc_2**, indicating land cover by colour:
```{r fig.width = 16, fig.height = 12.5}
DF.colour.scheme <- data.frame(
    land_cover = c("marsh",  "swamp",  "water",  "forest", "ag",     "shallow"),
    colour     = c("#000000","#E69F00","#56B4E9","#009E73","#F0E442","red"    )
    )
rownames(DF.colour.scheme) <- DF.colour.scheme[,"land_cover"]
my.palette <- DF.colour.scheme[,"colour"]

textsize.title <- 50
textsize.axis  <- 35

my.ggplot <- ggplot2::ggplot(data = NULL) + ggplot2::theme_bw();
my.ggplot <- my.ggplot + ggplot2::theme(
    title            = ggplot2::element_text(size = textsize.title, face = "bold"),
    axis.title.x     = ggplot2::element_text(size = textsize.axis,  face = "bold"),
    axis.title.y     = ggplot2::element_text(size = textsize.axis,  face = "bold"),
    axis.text.x      = ggplot2::element_text(size = textsize.axis,  face = "bold"),
    axis.text.y      = ggplot2::element_text(size = textsize.axis,  face = "bold"),
    strip.text.y     = ggplot2::element_text(size = textsize.axis,  face = "bold"),
    legend.title     = element_blank(),
    legend.text      = ggplot2::element_text(size = textsize.axis),
    panel.grid.major = ggplot2::element_line(colour = "gray", linetype = 2, size = 0.25),
    panel.grid.minor = ggplot2::element_line(colour = "gray", linetype = 2, size = 0.25)
    );

my.ggplot <- my.ggplot + ggplot2::labs(title = NULL, subtitle = NULL)
my.ggplot <- my.ggplot + ggplot2::scale_colour_manual(values = my.palette)
my.ggplot <- my.ggplot + ggplot2::scale_fill_manual(  values = my.palette)
my.ggplot <- my.ggplot + guides(
    colour = guide_legend(override.aes = list(alpha =  0.75, size = 5))
    )

my.ggplot <- my.ggplot + scale_x_continuous(limits = 300*c(-1,1), breaks = seq(-300,300,100))
my.ggplot <- my.ggplot + scale_y_continuous(limits = 150*c(-1,1), breaks = seq(-150,150, 50))

my.ggplot <- my.ggplot + ggplot2::xlab("FPC 1 score")
my.ggplot <- my.ggplot + ggplot2::ylab("FPC 2 score")

my.ggplot <- my.ggplot + geom_point(
    data    = DF.fpc,
    mapping = aes(x = fpc_1, y = fpc_2, colour = land_cover),
    size    = 0.5,
    alpha   = 0.5
    )

my.ggplot
```
We emphasize here that:

*   The functional principal component scores shown in the above scatter plot
    were computed for the time series in **DF.new**.

*   However, these scores (again, computed for **DF.new**) were computed
    with respect to the functional principal components
    extracted from time series in **DF.training**,
    which was used as the training data set.

# 5. Goodness-of-fit diagnostics of individual time series.

It is prudent to examine how well the B-spline and the FPC approximation
are fitting a given time series.

We demonstrate how to do so in this section.

The following code segment extracts, and then displays,
the 2017 VV time series of the very first-appearing location
in **DF.new**:
```{r}
temp.x_y <- DF.new[1,"your_x_y"]
print( temp.x_y );

DF.temp <- IW4[c("x","y","date","VV")];

DF.temp[,"year"] <- format(x = IW4[,'date'], format = "%Y")
is.selected.year <- ("2017" == DF.temp[,'year'])
DF.temp          <- DF.temp[is.selected.year,]

DF.temp[,"x_y"] <- apply(
    X      = DF.temp[,c('x','y')],
    MARGIN = 1,
    FUN    = function(x) { return(paste(x,collapse="_")) }
    )
colnames(DF.temp) <- paste0("their_",colnames(DF.temp))

DF.temp <- DF.temp[temp.x_y == DF.temp[,'their_x_y'],]
DF.temp <- DF.temp[,c("their_x_y","their_date","their_VV")]
knitr::kable(DF.temp, align = "c", row.names = FALSE)
```

We can examine the goodness-of-fit of the B-spline and FPC approximations
as follows:
```{r, fig.width = 16}
my.fpcFeatureEngine$plot.approximations(
    DF.input = DF.temp,
    location = "their_x_y",
    date     = "their_date",
    variable = "their_VV"
    )
```
Here, the black dots illustrate the given time series,
while the blue and red curves illustrate, respectively,
the fitted B-spline and the FPC approximation.
The two vertical dashed lines indicate the observation date range
of the training data.
