---
title: "FPC1 and FPC2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fpc1-fpc2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
    )
```
The R package **stcCropYield** implements a number of model training procedures
to build trained prediction models for early-season crop yield (parcel-level)
prediction for the province of Manitoba, using parcel-level NDVI and CARS-level
weather measurement time series as predictor variables.

This vignette demonstrates how to use the function
**stcCropYield::rollingWindowForwardValidation(.)**
to train a crop yield prediction model,
following the *rolling window forward validation* protocol,
using XGBoost as the underlying prediction technique.

First, load the **stcCropYield** R package.
```{r setup}
require(fpcFeatures)
require(ggplot2)
require(knitr)
require(logger)
```
Next, we generate a synthetic data set in the form of a single data frame
that has the structure expected by
**stcCropYield::rollingWindowForwardValidation(...)**:
We examine the structure of the data frame DF.training
```{r IW4}
data(IW4);
str(IW4);
```

Here are the first few rows of DF.training:
```{r headIW4}
knitr::kable(head(IW4), align = "c")
```

```{r}
DF.VV <- IW4[,c('x','y','date','VV')];
DF.VV[,'x_y'] <- base::apply(
    X      = DF.VV[,c('x','y')],
    MARGIN = 1,
    FUN    = function(x) { return(base::paste(x,collapse="_")) }
    );
```

```{r}
locations <- base::unique(DF.VV[,'x_y']);

set.seed(7654321);
is.subset.1 <- sample(x = c(TRUE,FALSE), size = length(locations), replace = TRUE);
locations.1 <- locations[ is.subset.1];
locations.2 <- locations[!is.subset.1];

DF.VV.1 <- DF.VV[DF.VV[,'x_y'] %in% locations.1,];
colnames(DF.VV.1) <- paste0("my_",colnames(DF.VV.1));

DF.VV.2 <- DF.VV[DF.VV[,'x_y'] %in% locations.2,];
colnames(DF.VV.2) <- paste0("your_",colnames(DF.VV.2));
```

```{r}
logger::log_threshold(level = logger::ERROR);

my.fpcFeatureEngine <- fpcFeatureEngine$new(
    training.data       = DF.VV.1,
    location            = 'my_x_y',
    date                = 'my_date',
    variable            = 'my_VV',
    n.partition         = 100,
    n.order             =   3,
    n.basis             =   9,
    smoothing.parameter =   0.1,
    n.harmonics         =   7
    );

my.fpcFeatureEngine$fit();
```

```{r harmonics, fig.height = 28, fig.width = 16}
my.fpcFeatureEngine$plot.harmonics();
```

```{r}
DF.land.cover <- base::unique(IW4[,c('x','y','land_cover')])
DF.land.cover[,'x_y'] <- base::apply(
    X      = DF.land.cover[,c('x','y')],
    MARGIN = 1,
    FUN    = function(x) { return(base::paste(x,collapse="_")) }
    );
DF.land.cover <- DF.land.cover[,c('x_y','land_cover')]
knitr::kable(head(DF.land.cover), align = "c", row.names = FALSE)
```

```{r transform}
DF.bspline.fpc <- my.fpcFeatureEngine$transform(
    newdata  = DF.VV.2,
    location = 'your_x_y',
    date     = 'your_date',
    variable = 'your_VV'
    );

DF.bspline.fpc <- base::merge(
    x    = DF.bspline.fpc,
    y    = DF.land.cover,
    by.x = 'your_x_y',
    by.y = 'x_y'
    )

DF.bspline.fpc <- DF.bspline.fpc[,c('your_x_y','year','land_cover',paste0('fpc_',1:7))]

knitr::kable(head(DF.bspline.fpc), align = "c", row.names = FALSE)
```

```{r fig.width = 16, fig.height = 12.5}
DF.colour.scheme <- data.frame(
    land_cover = c("marsh",  "swamp",  "water",  "forest", "ag",     "shallow"),
    colour     = c("#000000","#E69F00","#56B4E9","#009E73","#F0E442","red"    )
    )
rownames(DF.colour.scheme) <- DF.colour.scheme[,"land_cover"]
my.palette <- DF.colour.scheme[,"colour"]

textsize.title <- 50
textsize.axis  <- 35

my.ggplot <- ggplot2::ggplot(data = NULL) + ggplot2::theme_bw();
my.ggplot <- my.ggplot + ggplot2::theme(
    title            = ggplot2::element_text(size = textsize.title, face = "bold"),
    axis.title.x     = ggplot2::element_text(size = textsize.axis,  face = "bold"),
    axis.title.y     = ggplot2::element_text(size = textsize.axis,  face = "bold"),
    axis.text.x      = ggplot2::element_text(size = textsize.axis,  face = "bold"),
    axis.text.y      = ggplot2::element_text(size = textsize.axis,  face = "bold"),
    strip.text.y     = ggplot2::element_text(size = textsize.axis,  face = "bold"),
    legend.title     = element_blank(),
    legend.text      = ggplot2::element_text(size = textsize.axis),
    panel.grid.major = ggplot2::element_line(colour = "gray", linetype = 2, size = 0.25),
    panel.grid.minor = ggplot2::element_line(colour = "gray", linetype = 2, size = 0.25)
    );

my.ggplot <- my.ggplot + ggplot2::labs(title = NULL, subtitle = NULL)
my.ggplot <- my.ggplot + ggplot2::scale_colour_manual(values = my.palette)
my.ggplot <- my.ggplot + ggplot2::scale_fill_manual(  values = my.palette)
my.ggplot <- my.ggplot + guides(
    colour = guide_legend(override.aes = list(alpha =  0.75, size = 5))
    )

my.ggplot <- my.ggplot + scale_x_continuous(limits = 300*c(-1,1), breaks = seq(-300,300,100))
my.ggplot <- my.ggplot + scale_y_continuous(limits = 150*c(-1,1), breaks = seq(-150,150, 50))

my.ggplot <- my.ggplot + ggplot2::xlab("FPC 1 score")
my.ggplot <- my.ggplot + ggplot2::ylab("FPC 2 score")

my.ggplot <- my.ggplot + geom_point(
    data    = DF.bspline.fpc,
    mapping = aes(x = fpc_1, y = fpc_2, colour = land_cover),
    size    = 0.5,
    alpha   = 0.5
    )

my.ggplot
```

```{r}
temp.x_y <- DF.VV.2[1,"your_x_y"]
print( temp.x_y );

DF.temp <- IW4[c("x","y","date","VV")];

DF.temp[,"year"] <- format(x = IW4[,'date'], format = "%Y")
is.selected.year <- ("2017" == DF.temp[,'year'])
DF.temp          <- DF.temp[is.selected.year,]

DF.temp[,"x_y"] <- apply(
    X      = DF.temp[,c('x','y')],
    MARGIN = 1,
    FUN    = function(x) { return(paste(x,collapse="_")) }
    )
colnames(DF.temp) <- paste0("their_",colnames(DF.temp))

DF.temp <- DF.temp[temp.x_y == DF.temp[,'their_x_y'],]
DF.temp <- DF.temp[,c("their_x_y","their_date","their_VV")]
knitr::kable(DF.temp, align = "c", row.names = FALSE)
```

```{r, fig.width = 16}
my.fpcFeatureEngine$plot.approximations(
    DF.input = DF.temp,
    location = "their_x_y",
    date     = "their_date",
    variable = "their_VV"
    )
```
